---
title: "Data 607 Assignment 9 - Web APIs"
author: "Subhalaxmi Rout"
date: "3/29/2020"
output:
  pdf_document: default
  prettydoc::html_pretty: 
  html_document:
    theme: cayman
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##  Assignment overview

 The New York Times web site provides a rich set of APIs, as described here: https://developer.nytimes.com/apis </br>
You’ll need to start by signing up for an API key.

Your task is to choose one of the New York Times APIs, construct an interface in R to read in the JSON data, and transform it into an R DataFrame.

### Introduction

The aim of this assignment is the use of the New York Times web site API to create a data frame. 
 
 + Signed in New York Times
 + Registered for the API key
 + Chose `Article search` API
 + Set search keyword - `Coronavirus`

The result shows the key-word  `Coronavirus`  available article. 

### Load Library

```{r message=FALSE, warning=FALSE}
# loaded necessary library
library(jsonlite)
library(dplyr)
library(DT)
library(httr)
library(ggplot2)
```

### Load data

```{r}
# intialize api key
api_key <- "u5CCFWy8iVIWy5O86FPxnA93zWhJVlCY"

# initialize search word
search_word <- "Coronavirus"

# url for Article search api
url <- "https://api.nytimes.com/svc/search/v2/articlesearch.json?q="

# pass parameter such as search word and api_key
url <- paste0(url,search_word,"&api-key=",api_key,sep = "")

# store the value in a data frame
article <- jsonlite::fromJSON(url, flatten = TRUE) %>% data.frame()

# put data in to table format
DT::datatable(article, options = list(pageLength = 2))

```


### Select relevant column 
 In the data set we can see 33 columns, out of 33 columns I chose below columns. 
 
 + response.docs.abstract <- Abstract
 + response.docs.web_url <- URL
 + response.docs.source <- Source
 + response.docs.document_type <- DocType
 + response.docs.word_count <- WordCount
 + response.docs.headline.main <- HeadLine
 + response.docs.byline.original <- ByLine

 
```{r message=FALSE, warning=FALSE}
 df_article <- article %>% 
  dplyr::select (response.docs.abstract,response.docs.web_url
 , response.docs.source, response.docs.document_type, response.docs.word_count, 
 response.docs.headline.main, response.docs.byline.original) %>% 
  dplyr::rename("Abstract" = response.docs.abstract,"URL" = response.docs.web_url,
                  "Source" = response.docs.source, "DocType" = response.docs.document_type, 
                  "WordCount" = response.docs.word_count, "HeadLine" = response.docs.headline.main, 
                "Author" = response.docs.byline.original)

DT::datatable(df_article, options = list(pageLength = 2))
```

### Lonest article 

Author name and article wordcount is present in dataset. Here we find out which is the longest article on `Coronavirus`.

```{r, fig.width= 12, fig.height=10}
df_article_long <- data.frame(cbind(df_article$Author, df_article$WordCount))
# rename columns name
df_article_long <- df_article_long %>% dplyr::rename("Author" = X1, "WordCount" = X2 )
# change data type of wordcount column 
df_article_long$WordCount <- as.integer(as.character(df_article_long$WordCount))
# replace "BY" with ""
df_article_long$Author <- gsub( "By ", "",df_article_long$Author) 
# arrange in desc order
df_article_long <- df_article_long %>% arrange(desc(WordCount))

ggplot2::ggplot(data = df_article_long, aes(x = Author, y = WordCount)) + 
  geom_bar(aes(reorder(Author,WordCount),WordCount),stat = "identity",fill = "steelblue") + 
  coord_flip() + xlab("Author") + ylab("Word Count") +
  ggtitle("Longest article on Corona virus") +
  theme(plot.title = element_text(hjust = 0.5),panel.background = element_rect(fill = "white", color = NA)) +
  geom_text(aes( y = WordCount,label=WordCount), hjust = -0.20, color="Black", size=3.5)

```

###  Search key-word as a user input

This function will take the search word as user input and will return corresponding articles. 

Passed search word `politics`, user can pass any word, based on that word this function will return data.

```{r message=FALSE, warning=FALSE}
ArticleAPI <- function(word)
  
  {
  api_key <- "u5CCFWy8iVIWy5O86FPxnA93zWhJVlCY"
  
  url <-  paste("https://api.nytimes.com/svc/search/v2/articlesearch.json?q=",word,"&api-key=u5CCFWy8iVIWy5O86FPxnA93zWhJVlCY",sep="")
  
  paste0("Here is data for key-word: ", word)
  
  #get seems to fail sometimes
  df <- httr::RETRY("GET", url, times = 100, pause_base = 1)

  df <- content(df,as="text")
  
  df <-  as.data.frame(fromJSON(df))
  
  #clean up the column names
  colnames(df) <- gsub("^.*\\.","", colnames(df))
  
  return(df)
}

ArticleAPI("politics")

```

### Conclusion

This assignment taught us, how to using API to get data, manipulate data and do the analysis. Using `ArticleAPI` function users can get corresponding data present in the New York Times article search API.





